Probabilistic Symbolic Execution

### 2021-06-23 shared/Paper-Draft/PSE-POPL-2021.tex
WIP : Baseline Experiments

	\begin{algorithm}
		\caption{Getting Path Constraints : Symbolic Execution}
		\label{alg:symbex}
		\begin{algorithmic}[1]
			\Function{SymbEx}{$P_{prog} : Program$}
				\Let{$\varphi_{paths}$}{[]}, {$Ex_{stack}$} $\leftarrow$ {[]} \algorithmiccomment{List to store all the path conditions.}
				\Let{$I_{0}$}{\textsc{getStartInstruction}($P_{prog}$)}
				\Let{$S_{0}$}{[$I_{0}$, $\phi_{path}$, $\Delta$]} \algorithmiccomment{Empty Initial State}
				\State {$Ex_{stack}$.\textsc{Push}({$S_{0}$})} \algorithmiccomment{Start with $S_{cur}$ in Execution Stack}
				\While{$Ex_{stack}$ != $\phi$}
					\Let{$S_{cur}$}{$Ex_{stack}$.\textsc{Pop}( )}, $I_{cur} \leftarrow S_{cur}[1]$ \algorithmiccomment{Start State, [$I_{0}$, $\phi_{path}$, $\Delta$]}
					\Switch{\textsc{instType}($I_{cur}$)}
					\Case{v := e} \algorithmiccomment{Assignment Instruction}
						\Let{$I_{cur}$}{\textsc{getNextInstruction}($I_{cur}$)}
						\Let{$S_{cur}$}{[$I_{cur}$, $\phi_{path}$, $\Delta$[v $\rightarrow$ \textsc{Eval($e$, $\Delta$)}]]}
						\State {$Ex_{stack}$.\textsc{Push}({$S_{cur}$})}
					\EndCase
					\Case{\textbf{if} $(c_{sym})$ \textbf{then} $P_{1}$ \textbf{else} $P_{2}$} \algorithmiccomment{Branch Instruction} \label{line:branch}
						\If{(\textsc{isSat}($c_{sym}$) $\wedge$ \textsc{isSat}($\neg c_{sym}$))} \algorithmiccomment{Both cases SAT}
							\Let{$I_{1}$}{\textsc{getStartInstruction}($P_{1}$)} \Let{$I_{2}$} {\textsc{getStartInstruction}($P_{2}$)}	
							\Let{$S_{true}$}{[$I_{1}$, $\phi_{path} \wedge c_{sym}$, $\Delta$]}
							\Let{$S_{false}$}{[$I_{2}$, $\phi_{path} \wedge (\neg c_{sym})$, $\Delta$]}
							\State {$Ex_{stack}$.\textsc{Push}({$S_{false}$})}
							\State {$Ex_{stack}$.\textsc{Push}({$S_{true}$})} \algorithmiccomment{Start with True State}
						\ElsIf{\textsc{isSat}($c_{sym}$)} \algorithmiccomment{True case SAT}
							\Let{$I_{cur}$}{\textsc{getStartInstruction}($P_{1}$)}
							\Let {$S_{cur}$}{[$I_{cur}$, $\phi_{path} \wedge c_{sym}$, $\Delta$]}
							\State {$Ex_{stack}$.\textsc{Push}({$S_{cur}$})}
						\ElsIf{\textsc{isSat}($\neg c_{sym}$)} \algorithmiccomment{False case SAT}
							\Let{$I_{cur}$}{\textsc{getStartInstruction}($P_{2}$)}
							\Let{$S_{cur}$}{[$I_{2}$, $\phi_{path} \wedge (\neg c_{sym})$, $\Delta$]}
							\State {$Ex_{stack}$.\textsc{Push}({$S_{cur}$})}
						\Else \algorithmiccomment{$c_{sym}$ UNSAT}
							\Let{$I_{cur}$}{\textsc{getNextInstruction}($I_{cur}$)}
							\Let{$S_{cur}$}{[$I_{cur}$, $\phi_{path}$, $\Delta$]}
							\State {$Ex_{stack}$.\textsc{Push}({$S_{cur}$})}
						\EndIf
					\EndCase
					\Case{\textbf{assume}$(v == expr)$} \algorithmiccomment{Assume Instruction}
						\Let{$I_{cur}$}{\textsc{getNextInstruction}($I_{cur}$)}	
						\If{\textsc{isSat}($v == expr$)}
							\Let{$S_{cur}$}{[$I_{cur}$, $\phi_{path} \wedge (v == expr)$, $\Delta$]}
						\Else 
							\Let{$S_{cur}$}{[$I_{cur}$, $\phi_{path} \wedge $(\textsc{false}), $\Delta$]}
						\EndIf
						\State {$Ex_{stack}$.\textsc{Push}({$S_{cur}$})}
					\EndCase
					\Case{\textbf{HALT}} \algorithmiccomment{Terminate Instruction}
						\Let{$\varphi_{paths}$}{$\varphi_{paths}$.\textsc{Append}($\phi_{path}$)}		
					\EndCase
					\EndSwitch
				\EndWhile \\
				\Return{$\varphi_{paths}$}
			\EndFunction
		\end{algorithmic}
	\end{algorithm}

@@

	We use a robust symbolic execution engine, \textsc{KLEE} to generate the \textit{path constraints} corresponding to each path that our tool explores. For each path in the program, \textsc{KLEE} stores a list of \textit{constraints} that encodes the whole path. These constraints are over \textit{program} variables that have been marked \texttt{make\_symbolic()} by the user and are stored in \textsc{KLEE} as \textit{metadata} in the \textit{state} data structure during the \textit{dynamic} symbolic execution of the program.

	\textsc{KLEE} assigns values to these \textit{symbolic} variables by solving the current set of \textit{constraints} seen so far in the program path as described in Algorithm~\ref{alg:symbex} and also by concretizing some of the values that get evaluated as a result of \textit{concrete} execution. Upon reaching the \textit{branch} instruction at Line~\ref{line:branch} in Algorithm~\ref{alg:symbex}, \textsc{KLEE} forks the current \textit{state} by making two identical copies of it and then appending to each of \textit{new} states one additional constraint encoding the \textit{true} and \textit{false} side of the branch and adds it to the \textit{execution} stack containing \textit{states} that need to explored next. The process continues untill all the \textit{states} in the \textit{execution} stack are explored. Apart from \textit{path constraints}, the \textit{state} also contains a mapping of \textit{symbolic} variables to expressions that are either  \textit{symbolic} or \textit{concrete} (in the case of concrete execution of the program) and a list of \textit{instructions} currently getting executed as a part of the \textit{state}.

@@

	\section{Related Work}
	\label{sec:related}
	{\color{red} Subhajit \& Sumit: If you could both start looking into related work (Mayhap, original PSE paper, Axprof, P4WN, PSI), that would be great!}
	\begin{itemize}
		\item Compare and Contrast with Mayhap, PSE, AxProf, P4WN \& PSI.
		\item Explain how we differ by not using a Model Counting but a Constraint Solving approach.
		\item Support for \textsc{ForAll} variables along with \textit{probabilistic} variables. 
	\end{itemize}
	\section{Conclusion \& Future Work}
	\label{sec:conclusion}

>>>>

	\begin{algorithm}
		\caption{Getting Path Constraints : Symbolic Execution}
		\label{alg:symbex}
		\begin{algorithmic}[1]
			\Function{SymbEx}{$P_{prog} : Program$}
				\Let{$\varphi_{paths}$}{[]}, {$Ex_{stack}$} $\leftarrow$ {[]} \algorithmiccomment{List to store all the path conditions.}
				\Let{$I_{0}$}{\textsc{getStartInstruction}($P_{prog}$)}
				\Let{$S_{0}$}{[$I_{0}$, $\phi_{path}$, $\Delta$]} \algorithmiccomment{Empty Initial State}
				\State {$Ex_{stack}$.\textsc{Push}({$S_{0}$})} \algorithmiccomment{Start with $S_{cur}$ in Execution Stack}
				\While{$Ex_{stack}$ != $\phi$}
					\Let{$S_{cur}$}{$Ex_{stack}$.\textsc{Pop}( )}, $I_{cur} \leftarrow S_{cur}[1]$ \algorithmiccomment{Start State, [$I_{0}$, $\phi_{path}$, $\Delta$]}
					\Switch{\textsc{instType}($I_{cur}$)}
					\Case{v := e} \algorithmiccomment{Assignment Instruction}
						\Let{$I_{cur}$}{\textsc{getNextInstruction}($I_{cur}$)}
						\Let{$S_{cur}$}{[$I_{cur}$, $\phi_{path}$, $\Delta$[v $\rightarrow$ \textsc{Eval($e$, $\Delta$)}]]}
						\State {$Ex_{stack}$.\textsc{Push}({$S_{cur}$})}
					\EndCase
					\Case{\textbf{if} $(c_{sym})$ \textbf{then} $P_{1}$ \textbf{else} $P_{2}$} \algorithmiccomment{Branch Instruction} \label{line:branch}
						\If{(\textsc{isSat}($c_{sym}$) $\wedge$ \textsc{isSat}($\neg c_{sym}$))} \algorithmiccomment{Both cases SAT}
							\Let{$I_{1}$}{\textsc{getStartInstruction}($P_{1}$)} \Let{$I_{2}$} {\textsc{getStartInstruction}($P_{2}$)}	
							\Let{$S_{true}$}{[$I_{1}$, $\phi_{path} \wedge c_{sym}$, $\Delta$]}
							\Let{$S_{false}$}{[$I_{2}$, $\phi_{path} \wedge (\neg c_{sym})$, $\Delta$]}
							\State {$Ex_{stack}$.\textsc{Push}({$S_{false}$})}
							\State {$Ex_{stack}$.\textsc{Push}({$S_{true}$})} \algorithmiccomment{Start with True State}
						\ElsIf{\textsc{isSat}($c_{sym}$)} \algorithmiccomment{True case SAT}
							\Let{$I_{cur}$}{\textsc{getStartInstruction}($P_{1}$)}
							\Let {$S_{cur}$}{[$I_{cur}$, $\phi_{path} \wedge c_{sym}$, $\Delta$]}
							\State {$Ex_{stack}$.\textsc{Push}({$S_{cur}$})}
						\ElsIf{\textsc{isSat}($\neg c_{sym}$)} \algorithmiccomment{False case SAT}
							\Let{$I_{cur}$}{\textsc{getStartInstruction}($P_{2}$)}
							\Let{$S_{cur}$}{[$I_{2}$, $\phi_{path} \wedge (\neg c_{sym})$, $\Delta$]}
							\State {$Ex_{stack}$.\textsc{Push}({$S_{cur}$})}
						\Else \algorithmiccomment{Unconditional Branch}
							\Let{$I_{cur}$}{\textsc{getNextInstruction}($I_{cur}$)}
							\Let{$S_{cur}$}{[$I_{cur}$, $\phi_{path}$, $\Delta$]}
							\State {$Ex_{stack}$.\textsc{Push}({$S_{cur}$})}
						\EndIf
					\EndCase
					\Case{\textbf{assume}$(v == expr)$} \algorithmiccomment{Assume Instruction}
						\Let{$I_{cur}$}{\textsc{getNextInstruction}($I_{cur}$)}	
						\If{\textsc{isSat}($v == expr$)}
							\Let{$S_{cur}$}{[$I_{cur}$, $\phi_{path} \wedge (v == expr)$, $\Delta$]}
						\Else 
							\Let{$S_{cur}$}{[$I_{cur}$, $\phi_{path} \wedge $(\textsc{false}), $\Delta$]}
						\EndIf
						\State {$Ex_{stack}$.\textsc{Push}({$S_{cur}$})}
					\EndCase
					\Case{\textbf{HALT}} \algorithmiccomment{Terminate Instruction}
						\Let{$\varphi_{paths}$}{$\varphi_{paths}$.\textsc{Append}($\phi_{path}$)}		
					\EndCase
					\EndSwitch
				\EndWhile \\
				\Return{$\varphi_{paths}$}
			\EndFunction
		\end{algorithmic}
	\end{algorithm}

@@

	We use a robust dynamic symbolic execution engine, \textsc{KLEE} to generate the \textit{path constraints} corresponding to each path that our tool explores. For each path in the program, \textsc{KLEE} stores a list of \textit{constraints} that encodes the whole path. These constraints are over \textit{program} variables that have been marked \texttt{make\_symbolic()} by the user and are stored in \textsc{KLEE} as \textit{metadata} in the \textit{state} data structure during the \textit{dynamic} symbolic execution of the program.

	\textsc{KLEE} assigns values to these \textit{symbolic} variables by employing an \textsc{SMT Solver} for solving the current set of \textit{constraints} seen so far in the program path as described in Algorithm~\ref{alg:symbex} and also by concretizing some of the values that get evaluated as a result of \textit{concrete} execution (Eg. return value from external function calls). Upon reaching the \textit{branch} instruction at Line~\ref{line:branch} in Algorithm~\ref{alg:symbex}, \textsc{KLEE} forks the current \textit{state} by making two identical copies of it and then appending to each of \textit{new} states one additional constraint encoding the \textit{true} and \textit{false} side of the branch and adds it to the \textit{execution} stack containing \textit{states} that need to be explored next. The process continues untill all the \textit{states} in the \textit{execution} stack are explored. Apart from \textit{path constraints}, the \textit{state} also contains a mapping of \textit{symbolic} variables to expressions that are either  \textit{symbolic} or \textit{concrete} (in the case of concrete execution of the program) and a list of \textit{instructions} currently getting executed as a part of the \textit{state}.

@@

	\section{Related Work}
	\label{sec:related}
	{\color{red} Subhajit \& Sumit: If you could both start looking into related work (Mayhap, original PSE paper, Axprof, P4WN, PSI), that would be great!}
	\begin{itemize}
		\item Compare and Contrast with Mayhap, PSE, AxProf, P4WN \& PSI.
		\item Explain how we differ by not using Model Counting but a Constraint Solving approach.
		\item Support for \textsc{ForAll} variables along with \textit{probabilistic} variables. 
	\end{itemize}
	\section{Conclusion \& Future Work}
	\label{sec:conclusion}
